Assessing the impacts of time to detection distribution assumptions on detection probability estimation

### 2017-02-21 ReviewerComments.tex
Ongoing revisions.

I enjoyed reading this well-written manuscript, and believe that the modeling framework proposed is sufficiently versatile to be used in a wide range of applications. In addition to the points made by the referees, I have two comments:
\begin{enumerate}
\item Reviewer 1 is not convinced that it is helpful to make explicit the link to survival analysis. I tend to disagree with this, since I think that this is what renders the approach widely applicable and also extendable. In fact, the general idea strongly reminded me of several papers published recently by David Borchers and colleagues (using concepts from survival analysis to model detection events within distance sampling and spatially explicit capture-recapture). I was wondering if there is a common denominator between his work and the present manuscript (but the authors need not necessarily explore this).
\begin{quote}
We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
Their models are for removal-distance data, whereas ours is for removal-only data.
\end{quote}
\item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter. It was not clear to me why the rate parameter was chosen to be modeled as a function of covariates, as opposed to the mean of the TTDD. For example, in gamma GLMs, the mean is modeled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work). Basically I'm now wondering a) why the rate rather than the mean was modeled, and b) if that matters at all. In 3.1, a justification seems to be given, but I do not follow the reasoning here – why would the inclusion of random and fixed effects be any different if not the rate but the mean would be used? Sorry if I'm just missing the obvious.
\end{enumerate}

@@

As a potential user of such a model presented in this manuscript, I find the simulation analyses somewhat lacking.   
\begin{enumerate}
\item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
\begin{quote}
We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
However, it can be informative.
TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
\end{quote}
\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\begin{quote}
\adam{Investigate more/different simulations}
\end{quote}
\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\begin{quote}
\adam{We have not investigated either of these issues.
During my Masters work, we examined the effects of censoring on a mixture exponential model and found a 33\% increase in credible interval widths for $\pdet$ when switching from 9- to 3-interval data in a model with covariates.
However, I would not dream of extending those findings to other distributions.
We obviously would prefer longer durations, but those risk the violation of closure assumptions.
\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
I am inclined not to add Discussion on topics about which I do not have well-informed opinions.

To the reviewer's point, we can emphasize the value of a settling-in period toward justifying the use constant-rate assumptions in analysis.}
\end{quote}
\item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
\begin{quote}
\adam{Because of improvements in Stan, I want to run more simulation reps before quantifying this.}
\end{quote}
\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\begin{quote}
We have retained just the 50\% coverage.
\adam{Waiting until all new simulations are run.}
\end{quote}
\end{enumerate}

@@

Write self-contained paragraphs, with one main idea, not two or three (and each one half developed). As a reviewer, I enjoyed verifying that the cited papers in the introduction and in the discussion were indeed meaningful (by downloading them and reading those that I was not familiar with). However, when many papers are cited, it is useful to expand in the text why the different papers are being cited.  Doing so not only clarifies your intentions, but does proper justice to the papers being referred to. It is also a useful exercise because it allows you to tell whether you have more than one clear ideas and hence, material for more than one paragraph.  For instance, I think that the paragraph in the introduction, line 34, page 3, could be expanded and/or broken into two paragraphs.  One introducing the time-to-detection as is done in survival analysis, and one presenting to the reader how data that seem not to conform to the constant-detection assumption is dealt with by
using the idea of the increased detection probability via a mixture component. These are two different ideas, hence two different paragraphs. The same approach to construct paragraphs through the text should be taken.

>>>>

I enjoyed reading this well-written manuscript, and believe that the modeling framework proposed is sufficiently versatile to be used in a wide range of applications. In addition to the points made by the referees, I have two comments:
\begin{enumerate}
\item Reviewer 1 is not convinced that it is helpful to make explicit the link to survival analysis. I tend to disagree with this, since I think that this is what renders the approach widely applicable and also extendable. In fact, the general idea strongly reminded me of several papers published recently by David Borchers and colleagues (using concepts from survival analysis to model detection events within distance sampling and spatially explicit capture-recapture). I was wondering if there is a common denominator between his work and the present manuscript (but the authors need not necessarily explore this).
\begin{quote}
We agree, this manuscript has strong parallels to the continuous-time models of David Borchers and colleagues.  
Their models are for removal-distance data, whereas ours is for removal-only data.
\end{quote}
\item In Section 3.4, models are built where the various TTDDs considered are linked to covariates via the rate parameter. It was not clear to me why the rate parameter was chosen to be modeled as a function of covariates, as opposed to the mean of the TTDD. For example, in gamma GLMs, the mean is modeled as a function of covariates, with the shape parameter fixed (the latter as done also in the present work). Basically I'm now wondering a) why the rate rather than the mean was modeled, and b) if that matters at all. In 3.1, a justification seems to be given, but I do not follow the reasoning here – why would the inclusion of random and fixed effects be any different if not the rate but the mean would be used? Sorry if I'm just missing the obvious.
\begin{quote}
We have added a statement clarifying that there is no mathematical difference.
\adam{
In response to (a), while we fit our models with common distributions for $f_T(t)$, we could alternatively begin with a detection hazard rate $h(t)$.
Given this context, the detection rate feels more like the focus of inference than the expected time to detection.
}
\end{quote}
\end{enumerate}

@@

As a potential user of such a model presented in this manuscript, I find the simulation analyses somewhat lacking.   
\begin{enumerate}
\item Users will most certainly not bother using such complex modeling approach to model an intercept only - and thus, evaluating TTDDs based on intercept-only models seems overly simple and not very informative.
\begin{quote}
We agree that an intercept-only model is simplistic and will not be of much practical use in applied settings.
However, it can be informative.
TTDDs that are biased (e.g. non-mixture lognormal) or not robust to misspecification (e.g. mixture exponential) may not be trustworthy in more complicated analyses.
\end{quote}
\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\begin{quote}
\adam{Investigate more/different simulations}
\end{quote}
\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\begin{quote}
\adam{We have not investigated either of these issues.
During my Masters work, we examined the effects of censoring on a mixture exponential model and found a 33\% increase in credible interval widths for $\pdet$ when switching from 9- to 3-interval data in a model with covariates.
However, I would not dream of extending those findings to other distributions.
We obviously would prefer longer durations, but those risk the violation of closure assumptions.
\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
I am inclined not to add Discussion on topics about which I do not have well-informed opinions.
}
We have added a mention of the trade-offs of a settling down period to reduce observer effects on behavior.
\end{quote}

\item Providing computation times for scenarios that vary by model complexity and sample size (including the computer used for analyses) may also provide a gauge for users as to whether this method is feasible for their needs.
\begin{quote}
\adam{Because of improvements in Stan, I want to run more simulation reps before quantifying this.}
\end{quote}
\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\begin{quote}
We have retained just the 50\% coverage.
\adam{Waiting until all new simulations are run.}
\end{quote}
\end{enumerate}

@@

Write self-contained paragraphs, with one main idea, not two or three (and each one half developed). As a reviewer, I enjoyed verifying that the cited papers in the introduction and in the discussion were indeed meaningful (by downloading them and reading those that I was not familiar with). However, when many papers are cited, it is useful to expand in the text why the different papers are being cited.  Doing so not only clarifies your intentions, but does proper justice to the papers being referred to. It is also a useful exercise because it allows you to tell whether you have more than one clear ideas and hence, material for more than one paragraph.  For instance, I think that the paragraph in the introduction, line 34, page 3, could be expanded and/or broken into two paragraphs.  One introducing the time-to-detection as is done in survival analysis, and one presenting to the reader how data that seem not to conform to the constant-detection assumption is dealt with by
using the idea of the increased detection probability via a mixture component. These are two different ideas, hence two different paragraphs. The same approach to construct paragraphs through the text should be taken.
\begin{quote}
We have reviewed the entire manuscript with these points in mind.
We have split the cited paragraph and others.
\adam{The only times I feel I have cited many articles is when I've said ``Many authors have done this.'.}
\end{quote}
